---
title: "Phonation Feature Analysis of mPower voice data V1"
author: "Thanneer Perumal"
date: "`r date()`"
output: html_document
---
```{r libs, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE}
## It is assumed your working directory is where this file is

# Clear R console screen output
cat("\014")  

# Load required libraries
library('synapseClient')
library('knitr')
library('dplyr')
library('data.table')
library('rjson')
library('plyr')
library('stringr')
library('rGithubClient')
library('devtools')
library('ggplot2')
library('reshape2')
library('psych')
library('fpc')
library('RColorBrewer')
library('knit2synapse')

# knit2synapse::storeAndKnitToFileEntity('./FeatureAnalysis.Rmd', parentId = "syn4557550", entityName = 'Feature Analysis')

## Requires ggdendro
devtools::source_url("https://raw.githubusercontent.com/chr1swallace/random-functions/master/R/ggplot-heatmap.R")

synapseLogin()
# synapseCacheDir('/mPower')
paretnId = 'syn4557550'

# source utility files from ../R/ folder
file.sources = list.files('../R/lib/',pattern="*")
file.sources = sapply(file.sources,function(x){return(paste('../R/lib',x,sep='/'))})
tmp = sapply(file.sources,source,.GlobalEnv)

options(xtable.type="html")

knitr::opts_chunk$set(
  echo=FALSE,
  warning=FALSE,
  message=FALSE,
  error = FALSE,
  tidy = FALSE,
  cache = TRUE)
```
### Voice metadata table download
```{r voice.table.download}
VOICE_ID = 'syn4726961'
VOICE = fread(synGet(VOICE_ID)@filePath, data.table=F); 

SURVEY_ID = 'syn4726964'
SURVEY = fread(synGet(SURVEY_ID)@filePath, data.table=F)

UPDRS_ID = 'syn4726968'
UPDRS = fread(synGet(UPDRS_ID)@filePath, data.table=F)
```
### Merge voice and survey
```{r merge.voice.survey}
# Merge voice and survey data
VOICE.SURVEY = merge(VOICE,SURVEY, by = 'healthCode', all=F)

# Remove in complete audio and countdown recordings
VOICE.SURVEY = VOICE.SURVEY[complete.cases(VOICE.SURVEY[,grep('audio\\.',colnames(VOICE.SURVEY),value=T)]) &
                              complete.cases(VOICE.SURVEY[,grep('countdown\\.',colnames(VOICE.SURVEY),value=T)]),]
colnames(VOICE.SURVEY) = sapply(colnames(VOICE.SURVEY), str_replace_all,'[^[:alnum:]]', '.')
```
```{r var.of.int}
# Covariates
FactCovars =  c("appVersion.x", "phoneInfo.x", "momentInDayFormat.json.choiceAnswers", "calculatedMeds",
                "are.caretaker", "deep.brain.stimulation", "diagnosis.year", "education", "employment", "gender", "healthcare.provider",
                "maritalStatus", "medical.usage", "medical.usage.yesterday", "professional.diagnosis", "race",  
                "smartphone", "smoked", "surgery")
ContCovars =   c("onset.year", "age", "years.smoking")
DateVars = c("createdOn.x")
# Feature Sets
AudioFeatures = c("audio.Median.F0", "audio.Mean.Jitter", "audio.Median.Jitter", "audio.Mean.Shimmer", "audio.Median.Shimmer",
                "audio.MFCC.Band.1", "audio.MFCC.band.2", "audio.MFCC.band.3", "audio.MFCC.band.4", "audio.MFCC.Jitter.band.1.Positive",
                "audio.MFCC.Jitter.band.2.Positive", "audio.MFCC.Jitter.band.3.Positive", "audio.MFCC.Jitter.band.4.Positive")
CountdownFeatures = c("countdown.Median.F0", "countdown.Mean.Jitter", "countdown.Median.Jitter", "countdown.Mean.Shimmer",
                "countdown.Median.Shimmer", "countdown.MFCC.Band.1", "countdown.MFCC.band.2", "countdown.MFCC.band.3",
                "countdown.MFCC.band.4", "countdown.MFCC.Jitter.band.1.Positive", "countdown.MFCC.Jitter.band.2.Positive",
                "countdown.MFCC.Jitter.band.3.Positive", "countdown.MFCC.Jitter.band.4.Positive")
```

### Preprocess data
* Retain recordings with 'version 1.0.5, build 12' and 'version 1.0, build 7'
* Retain samples with non NA "professional-daignosis"
* Retain samples with age less than 100
* Retain samples that are only less than current date
```{r filtering, echo=TRUE}
VOICE.SURVEY <- 
  VOICE.SURVEY %>%
  filter(appVersion.x == 'version 1.0.5, build 12' | appVersion.x == 'version 1.0, build 7') %>%
  filter(!is.na(professional.diagnosis)) %>%
  filter(age <= 100) %>%
  filter(as.Date(createdOn.x) <= Sys.Date())
```
`r length(REMOVED_RECORDS)` samples were removed from analysis

```{r data.diagnosis, fig.height=10, fig.width=10}
# Plot number of case-control recordings
tmp = split(VOICE.SURVEY,as.Date(VOICE.SURVEY$createdOn.x))
nCaseControl = cbind(sapply(tmp, function(x){return(dim(x)[1])}),
                     t(sapply(tmp, function(x){summary(factor(x$"professional.diagnosis"))})))
colnames(nCaseControl) = c('Total','Control','Case')
nCaseControl = rownameToFirstColumn(nCaseControl, 'Date')
nCaseControl = melt(nCaseControl, 'Date')

p = ggplot(nCaseControl, aes(x=factor(Date), y = value, group = factor(variable), color= factor(variable))) + geom_line()
p = p + theme(axis.text.x = element_text(angle = 90, hjust = 1), text = element_text(size=18)) + xlab("Date") + ylab("Count") +  ggtitle("Number of case - control")
p

# Number of recordings per participant
tmp = split(VOICE.SURVEY, factor(VOICE.SURVEY$healthCode))
len.rec = sapply(tmp, function(x){dim(x)[1]})
len.date = sapply(tmp, function(x){length(unique(x$uploadDate.x))})

H = data.frame(healthCode = names(tmp), 
               nRecordings = sapply(tmp, function(x){return(dim(x)[1])}),
               Dx = ifelse(sapply(tmp, function(x){return(unique(x$professional.diagnosis))}) == 'false',
                           'Control','Case'))
p <- ggplot(H,aes(nRecordings, fill=Dx))+geom_density()
p

## Correlation between covariates and Dx
COVARS = VOICE.SURVEY[,c(FactCovars,ContCovars)]
rownames(COVARS) = VOICE.SURVEY$rowID
  
COVARS[,FactCovars] = lapply(COVARS[,FactCovars], factor)
COVARS[,ContCovars] = lapply(COVARS[,ContCovars], as.numeric)

g <- getAssociationStatistics(COVARS[1:1000,])
ggheatmap.show(g$plot, col.width=0.3, row.width=0.15)
```
### PCA of phonation features
```{r decompse.phonation.data, fig.height=8, fig.width=8}
# Find principal components of audio phonation features
PHONATION = t(as.matrix(VOICE.SURVEY[,AudioFeatures]))
colnames(PHONATION) = VOICE.SURVEY$rowID
PC <- prcomp(PHONATION[,1:1000])

# Determine number of clusters automatically using pamk
pam.cluster <- fpc::pamk(t(PHONATION[,1:1000]), krange=2:10)

# Plot first 2 PCs
plotdata <- data.frame(UID=rownames(PC$rotation), 
                       PC1=PC$rotation[,1], 
                       PC2=PC$rotation[,2], 
                       cluster=pam.cluster$pamobject$clustering)

plotdata <- cbind(plotdata, COVARS[1:1000,])
plotdata <- mutate(plotdata, labels = '');#ifelse(Diffname_short == '',as.character(UID),'')) 

p <- ggplot(plotdata, aes(x=PC1, y=PC2))
p <- p + geom_point(aes(color=factor(cluster), shape=professional.diagnosis, size=gender))
p <- p + theme_bw() + scale_size_manual(values = c(4, 2)) + scale_shape_manual(values=1:2) + theme(legend.position="top") 
p <- p + geom_text(aes(label= labels), size=4, hjust=0)
p
```

### Remove outliers
* Retain recordings with healthCodes having atleast more than 2 recordings 
```{r filtering, echo=TRUE}
VOICE.SURVEY <- 
  VOICE.SURVEY %>%
  filter(healthCode %in% names(len.rec)[which(len.rec>2)]) %>%
  filter(healthCode %in% names(len.date)[which(len.date>2)])
```

```{r data.diagnosis, fig.height=10, fig.width=10}
# Plot number of case-control recordings
tmp = split(VOICE.SURVEY,as.Date(VOICE.SURVEY$createdOn.x))
nCaseControl = cbind(sapply(tmp, function(x){return(dim(x)[1])}),
                     t(sapply(tmp, function(x){summary(factor(x$"professional.diagnosis"))})))
colnames(nCaseControl) = c('Total','Control','Case')
nCaseControl = rownameToFirstColumn(nCaseControl, 'Date')
nCaseControl = melt(nCaseControl, 'Date')

p = ggplot(nCaseControl, aes(x=factor(Date), y = value, group = factor(variable), color= factor(variable))) + geom_line()
p = p + theme(axis.text.x = element_text(angle = 90, hjust = 1), text = element_text(size=18)) + xlab("Date") + ylab("Count") +  ggtitle("Number of case - control")
p

# Number of recordings per participant
tmp = split(VOICE.SURVEY, factor(VOICE.SURVEY$healthCode))
H = data.frame(healthCode = names(tmp), 
               nRecordings = sapply(tmp, function(x){return(dim(x)[1])}),
               Dx = ifelse(sapply(tmp, function(x){return(unique(x$professional.diagnosis))}) == 'false',
                           'Control','Case'))
p <- ggplot(H,aes(nRecordings, fill=Dx))+geom_density()
p

## Correlation between covariates and Dx
COVARS = VOICE.SURVEY[,c(FactCovars,ContCovars)]
rownames(COVARS) = VOICE.SURVEY$rowID
  
COVARS[,FactCovars] = lapply(COVARS[,FactCovars], factor)
COVARS[,ContCovars] = lapply(COVARS[,ContCovars], as.numeric)

g <- getAssociationStatistics(COVARS[1:1000,])
ggheatmap.show(g$plot, col.width=0.3, row.width=0.15)
```

```{r decompse.phonation.data, fig.height=8, fig.width=8}
# Find principal components of audio phonation features
PHONATION = t(as.matrix(VOICE.SURVEY[,AudioFeatures]))
colnames(PHONATION) = VOICE.SURVEY$rowID
PC <- prcomp(PHONATION[,1:1000])

# Determine number of clusters automatically using pamk
pam.cluster <- fpc::pamk(t(PHONATION[,1:1000]), krange=2:10)

# Plot first 2 PCs
plotdata <- data.frame(UID=rownames(PC$rotation), 
                       PC1=PC$rotation[,1], 
                       PC2=PC$rotation[,2], 
                       cluster=pam.cluster$pamobject$clustering)

plotdata <- cbind(plotdata, COVARS[1:1000,])
plotdata <- mutate(plotdata, labels = '');#ifelse(Diffname_short == '',as.character(UID),'')) 

p <- ggplot(plotdata, aes(x=PC1, y=PC2))
p <- p + geom_point(aes(color=factor(cluster), shape=professional.diagnosis, size=gender))
p <- p + theme_bw() + scale_size_manual(values = c(4, 2)) + scale_shape_manual(values=1:2) + theme(legend.position="top") 
p <- p + geom_text(aes(label= labels), size=4, hjust=0)
p
```

### Store in synapse
```{r synapsestore, echo=FALSE, include=FALSE, eval=TRUE}
ActivityName <- 'Phonation Feature Analysis'
  
thisFileName <- 'FeatureAnalysis.Rmd'
  
# Github link
thisRepo <- getRepo(repository = "th1vairam/mPower_Data_Analysis", 
                    ref="branch", 
                    refName='master')

thisFile <- getPermlink(repository = thisRepo,
                        repositoryPath=paste0('code/Rmd/', thisFileName))
    
# Populate wiki with results
CODE <- Folder(name = 'Feature Analysis',parentId = parentId)
CODE <- synStore(CODE)
```