---
title: "Classifying UPDRS scores based on phonation features"
author: "Thanneer Perumal"
date: "`r date()`"
output: html_document
---
```{r libs, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE}
## It is assumed your working directory is where this file is

# Clear R console screen output
cat("\014")  

# Load required libraries
library('synapseClient')
library('knitr')
library('dplyr')
library('data.table')
library('rjson')
library('plyr')
library('stringr')
library('rGithubClient')
library('devtools')
library('ggplot2')
library('reshape2')
library('psych')
library('fpc')
library('RColorBrewer')
library('knit2synapse')

# knit2synapse::storeAndKnitToFileEntity('./UPDRS_Analysis.Rmd', entityName = 'UPDRS Analysis', parentId = "syn4557550")

## Requires ggdendro
devtools::source_url("https://raw.githubusercontent.com/chr1swallace/random-functions/master/R/ggplot-heatmap.R")

synapseLogin()
# synapseCacheDir('/mPower')
parentId = 'syn4557550'

# source utility files from ../R/ folder
file.sources = list.files('../R/lib/',pattern="*")
file.sources = sapply(file.sources,function(x){return(paste('../R/lib',x,sep='/'))})
tmp = sapply(file.sources,source,.GlobalEnv)

options(xtable.type="html")

knitr::opts_chunk$set(
  echo=FALSE,
  warning=FALSE,
  message=FALSE,
  error = FALSE,
  tidy = FALSE,
  cache = TRUE)
```

### Voice metadata table download
```{r voice.table.download}
VOICE_ID = 'syn4726961'
ALL_USED_IDs = VOICE_ID
VOICE = fread(synGet(VOICE_ID)@filePath, data.table=F); 

SURVEY_ID = 'syn4726964'
ALL_USED_IDs = c(ALL_USED_IDs, SURVEY_ID)
SURVEY = fread(synGet(SURVEY_ID)@filePath, data.table=F)

UPDRS_ID = 'syn4726968'
ALL_USED_IDs = c(ALL_USED_IDs, UPDRS_ID)
UPDRS = fread(synGet(UPDRS_ID)@filePath, data.table=F)
```
### Merge voice and survey
```{r merge.voice.survey}
# Merge voice and survey data
setnames(VOICE, c("uploadDate","createdOn"), paste(c("uploadDate","createdOn"),'voice',sep='.')) 
setnames(SURVEY, c("uploadDate","createdOn"), paste(c("uploadDate","createdOn"),'survey',sep='.')) 
VOICE.SURVEY = merge(VOICE,SURVEY, by = c('healthCode','appVersion','phoneInfo'), all=F)

# Remove in complete audio and countdown recordings
VOICE.SURVEY = VOICE.SURVEY[complete.cases(VOICE.SURVEY[,grep('audio\\.',colnames(VOICE.SURVEY),value=T)]) &
                              complete.cases(VOICE.SURVEY[,grep('countdown\\.',colnames(VOICE.SURVEY),value=T)]),]
colnames(VOICE.SURVEY) = sapply(colnames(VOICE.SURVEY), str_replace_all,'[^[:alnum:]]', '.')
```
### Merge voice, survey and updrs data
```{r merge.voice.survey.updrs}
# Create data variables to merge
VOICE.SURVEY <- mutate(VOICE.SURVEY, createdOn.voice = as.POSIXct(createdOn.voice), createdDate = as.Date(createdOn.voice))
setnames(UPDRS, c("uploadDate","createdOn"), paste(c("uploadDate","createdOn"),'updrs',sep='.')) 
UPDRS <- mutate(UPDRS, createdOn.updrs = as.POSIXct(createdOn.updrs), createdDate = as.Date(createdOn.updrs))

# Subset data that have both voice and updrs in at least one time
healthCode = intersect(UPDRS$healthCode, VOICE.SURVEY$healthCode)
VOICE.SURVEY <- VOICE.SURVEY[VOICE.SURVEY$healthCode %in% healthCode,]
UPDRS <- UPDRS[UPDRS$healthCode %in% healthCode,]

# Split updrs and voice.survey per participants
VOICE.SURVEY.PP <- split(VOICE.SURVEY, factor(VOICE.SURVEY$healthCode))
UPDRS.PP <- split(UPDRS, factor(UPDRS$healthCode))

# Merge UPDRS table to voice survey table
VOICE.SURVEY.UPDRS.PP = mapply(function(updrs,voice.survey){
  tmp1 = merge(updrs,voice.survey, by = c("createdDate", "healthCode",  "appVersion",  "phoneInfo"), all = T)
  return(list(tmp1))
},
             UPDRS.PP,
             VOICE.SURVEY.PP)

VOICE.SURVEY.UPDRS = data.frame(rbindlist(VOICE.SURVEY.UPDRS.PP))
```

```{r var.of.int}
# Covariates
FactCovars =  c("appVersion", "phoneInfo", "EQ.5D1", "GELTQ.1a", "GELTQ.1b", "GELTQ.1c", "GELTQ.2", "MDS.UPRDRS1.1",  "MDS.UPDRS1.3", 
                "MDS.UPDRS1.4", "MDS.UPDRS1.5", "MDS.UPDRS1.7", "MDS.UPDRS1.8", "MDS.UPDRS2.1", "MDS.UPDRS2.4", "MDS.UPDRS2.5", "MDS.UPDRS2.6", "MDS.UPDRS2.7",
                "MDS.UPDRS2.8", "MDS.UPDRS2.9", "MDS.UPDRS2.10", "MDS.UPDRS2.12", "MDS.UPDRS2.13", "momentInDayFormat.json.choiceAnswers", 
                "momentInDayFormat.json.questionTypeName", "calculatedMeds", "are.caretaker", "deep.brain.stimulation", "education", "employment", "gender",
                "healthcare.provider", "maritalStatus",  "medical.usage", "medical.usage.yesterday", "professional.diagnosis", "race", "smartphone", "smoked",
                "surgery")
ContCovars =   c("onset.year", "diagnosis.year", "age", "years.smoking")
DateVars = c("createdDate", "uploadDate.updrs", "createdOn.updrs", "createdOn.voice", "uploadDate.voice",  "uploadDate.survey", "createdOn.survey", "last.smoked")

# Feature Sets
updrsFeatures = c("MDS.UPRDRS1.1",  "MDS.UPDRS1.3", "MDS.UPDRS1.4", "MDS.UPDRS1.5", "MDS.UPDRS1.7", "MDS.UPDRS1.8", "MDS.UPDRS2.1", "MDS.UPDRS2.4", 
                 "MDS.UPDRS2.5", "MDS.UPDRS2.6", "MDS.UPDRS2.7", "MDS.UPDRS2.8", "MDS.UPDRS2.9", "MDS.UPDRS2.10", "MDS.UPDRS2.12", "MDS.UPDRS2.13")
AudioFeatures = c("audio.Median.F0", "audio.Mean.Jitter", "audio.Median.Jitter", "audio.Mean.Shimmer", "audio.Median.Shimmer",
                "audio.MFCC.Band.1", "audio.MFCC.band.2", "audio.MFCC.band.3", "audio.MFCC.band.4", "audio.MFCC.Jitter.band.1.Positive",
                "audio.MFCC.Jitter.band.2.Positive", "audio.MFCC.Jitter.band.3.Positive", "audio.MFCC.Jitter.band.4.Positive")
CountdownFeatures = c("countdown.Median.F0", "countdown.Mean.Jitter", "countdown.Median.Jitter", "countdown.Mean.Shimmer",
                "countdown.Median.Shimmer", "countdown.MFCC.Band.1", "countdown.MFCC.band.2", "countdown.MFCC.band.3",
                "countdown.MFCC.band.4", "countdown.MFCC.Jitter.band.1.Positive", "countdown.MFCC.Jitter.band.2.Positive",
                "countdown.MFCC.Jitter.band.3.Positive", "countdown.MFCC.Jitter.band.4.Positive")
```

### Preprocess 
* Retain recordings with 'version 1.0.5, build 12' and 'version 1.0, build 7'
* Retain samples with non NA "professional-daignosis"
* Retain samples with age less than 100 and non NA age
* Retain samples that are only less than current date
```{r filtering, echo=TRUE}
VOICE.SURVEY.UPDRS = VOICE.SURVEY.UPDRS %>%
  mutate(row_ID = 1:dim(VOICE.SURVEY.UPDRS)[1])
tmp = na.omit(VOICE.SURVEY.UPDRS[,c('row_ID',updrsFeatures,AudioFeatures,CountdownFeatures)])
             
VOICE.SURVEY.UPDRS <- 
  VOICE.SURVEY.UPDRS %>%
  filter(row_ID %in% tmp$row_ID) %>%
  filter(appVersion == 'version 1.0.5, build 12' | appVersion == 'version 1.0, build 7') %>%
  filter(!is.na(professional.diagnosis)) %>%
  filter(age <= 100 | !is.na(age)) %>%
  filter(as.Date(createdDate) <= Sys.Date())

VOICE.SURVEY.UPDRS
```

### Distribution of samples
```{r data.diagnosis}
# Plot number of case-control recordings
VOICE.SURVEY.UPDRS.PerDate = split(VOICE.SURVEY.UPDRS,as.Date(VOICE.SURVEY.UPDRS$createdDate))
VOICE.SURVEY.UPDRS.PerDate.PP = lapply(VOICE.SURVEY.UPDRS.PerDate, function(x){ split(x,x$healthCode) })

nCaseControl = t(sapply(VOICE.SURVEY.UPDRS.PerDate.PP, function(x){  
  tmp = sapply(x,function(x){ unique(x$professional.diagnosis) })  
  return(c(total = length(x), control = sum(tmp == 'false'), case = sum(tmp == 'true')))
}))
colnames(nCaseControl) = c('Total','Control','Case')
nCaseControl = rownameToFirstColumn(nCaseControl, 'Date')
nCaseControl = melt(nCaseControl, 'Date')

p = ggplot(nCaseControl, aes(x=factor(Date), y = value, group = factor(variable), color= factor(variable))) + geom_line()
p = p + theme(axis.text.x = element_text(angle = 90, hjust = 1), text = element_text(size=18)) + xlab("Date") + ylab("Count") +  ggtitle("Number of case - control that have all voice, survey and updrs")
p

# Number of recordings per participant
VOICE.SURVEY.UPDRS.PP = split(VOICE.SURVEY.UPDRS, factor(VOICE.SURVEY.UPDRS$healthCode))
len.rec = sapply(VOICE.SURVEY.UPDRS.PP, function(x){dim(x)[1]})
len.date = sapply(VOICE.SURVEY.UPDRS.PP, function(x){length(unique(x$uploadDate.voice))})

H = data.frame(healthCode = names(VOICE.SURVEY.UPDRS.PP), 
               nRecordings = sapply(VOICE.SURVEY.UPDRS.PP, function(x){return(dim(x)[1])}),
               Dx = ifelse(sapply(VOICE.SURVEY.UPDRS.PP, function(x){return(unique(x$professional.diagnosis))}) == 'false',
                           'Control','Case'))
p <- ggplot(H,aes(nRecordings, fill=Dx)) + geom_density() +  ggtitle("PDF of recordings that have all voice, survey and updrs")
p

## Correlation between covariates and Dx
COVARS = VOICE.SURVEY.UPDRS[,c(FactCovars,ContCovars)]
COVARS = unique(COVARS)
  
writeLines('Correlation between Factor Variables')
COVARS[,FactCovars] = lapply(COVARS[,FactCovars,drop=F], factor)
g <- getAssociationStatistics(COVARS[,FactCovars])
ggheatmap.show(g$plot, col.width=0.3, row.width=0.15)

#COVARS[,ContCovars] = lapply(COVARS[,ContCovars,drop=F], as.numeric)
writeLines('Correlation between Factor and Continuous Variables')
COVARS[,c(FactCovars, ContCovars)] = lapply(COVARS[,c(FactCovars, ContCovars),drop=F], as.numeric)
g <- getAssociationStatistics(COVARS[,c(FactCovars, ContCovars)])
ggheatmap.show(g$plot, col.width=0.3, row.width=0.15)
```

### Effect of age
```{r age.related.effects, fig.height=15, fig.width=15}
VOICE.SURVEY.UPDRS[,FactCovars] = lapply(VOICE.SURVEY.UPDRS[,FactCovars], factor)
VOICE.SURVEY.UPDRS[,ContCovars] = lapply(VOICE.SURVEY.UPDRS[,ContCovars], as.numeric)

p <-list()
for (i in updrsFeatures){
  p[[i]] = ggplot(VOICE.SURVEY.UPDRS, aes_string(x = i, y = "age", fill = "professional.diagnosis")) +  geom_boxplot()
}
multiplot(plotlist=p,cols=4)
```

### PCA of phonation features
```{r decompse.phonation.data, fig.height=20, fig.width=20}
# Find principal components of audio phonation features
PHONATION = t(as.matrix(VOICE.SURVEY.UPDRS[,AudioFeatures]))
colnames(PHONATION) = 1:dim(PHONATION)[2]
PC <- prcomp(PHONATION)

# Plot all PCs
p1 <- list()
for (i in 1:5)
  for (j in 1:5)
    if (i == j){
      plotdata = cbind(1:13,as.data.frame(PC$sdev))
      colnames(plotdata) = c('PC','SDEV')
      p <- ggplot(plotdata, aes(x=PC, y = SDEV)) + geom_bar(stat='identity')
      p1 <- c(p1, list(p))
    } else {
      plotdata <- data.frame(UID=rownames(PC$rotation),
                             PC1 = PC$rotation[,i],
                             PC2 = PC$rotation[,j])
      
      plotdata <- cbind(plotdata, VOICE.SURVEY.UPDRS)
      plotdata <- mutate(plotdata, labels = '');#ifelse(Diffname_short == '',as.character(UID),'')) 
      
      p <- ggplot(plotdata, aes(x=PC1, y=PC2))
      p <- p + geom_point(aes(color=factor(professional.diagnosis), shape = factor(gender), size = factor(medical.usage)))
      p <- p + theme_bw() + scale_size_manual(values = c(8,6,4,2)) + scale_shape_manual(values=1:3) + theme(legend.position="top") 
      p <- p + geom_text(aes(label= labels), size=4, hjust=0)
      p1 <- c(p1, list(p))      
    }
multiplot(plotlist=p1,cols=5)  
```

### Store in synapse
```{r synapsestore, echo=FALSE, include=FALSE, eval=TRUE}
ActivityName <- 'UPDRS Analysis With Phonation Feature Analysis'
  
thisFileName <- 'UPDRS_Analysis.Rmd'
  
# Github link
thisRepo <- getRepo(repository = "th1vairam/mPower_Data_Analysis", 
                    ref="branch", 
                    refName='master')

thisFile <- getPermlink(repository = thisRepo,
                        repositoryPath=paste0('code/Rmd/', thisFileName))
    
# Populate wiki with results
CODE <- File('./UPDRS_Analysis.Rmd',name = 'UPDRS Analysis',parentId = parentId)
CODE <- synStore(CODE, executed = thisFile, used = ALL_USED_IDs, activityName = ActivityName)
```