---
title: "Quality control analysis and feature extraction of mPower voice data V1"
author: "Thanneer Perumal"
date: "`r date()`"
output: html_document
---
```{r libs, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE}
## It is assumed your working directory is where this file is
 
# Clear R console screen output
cat("\014")  

# Load required libraries
library('synapseClient')
library('knitr')
library('dplyr')
library('data.table')
library('rjson')
library('plyr')
library('stringr')
library('rGithubClient')
library('devtools')
library('knit2synapse')

# devtools::install_github(repo = "knit2synapse-1", username = "th1vairam", ref="dev")
# library(knit2synapse)
# storeAndKnitToFileEntity('./VoiceDataDownload.Rmd', parentId = "syn4557550", entityName = "Phonation Feature Extraction and Quality Control")

synapseLogin()
synapseCacheDir('/mPower')

# source utility files from ../R/ folder
file.sources = list.files('../R/',pattern="get*")
file.sources = sapply(file.sources,function(x){return(paste('../R',x,sep='/'))})
tmp = sapply(file.sources,source,.GlobalEnv)

options(xtable.type="html")

knitr::opts_chunk$set(
  echo=FALSE,
  warning=FALSE,
  message=FALSE,
  error = FALSE,
  tidy = FALSE,
  cache = FALSE)
```

### Voice metadata table download
```{r voice.table.download}
VOICE_ID = 'syn4590865';#'syn3420254'
ALL_USED_IDs = VOICE_ID
VOICE_TBL = synTableQuery(paste("SELECT * FROM",VOICE_ID))
VOICE = VOICE_TBL@values
VOICE$rowID = rownames(VOICE)

SURVEY_ID = 'syn4590867';#'syn3474927'
ALL_USED_IDs = c(ALL_USED_IDs, SURVEY_ID)
SURVEY_TBL = synTableQuery(paste('SELECT * FROM',SURVEY_ID)) 
SURVEY = SURVEY_TBL@values

UPDRS_ID = 'syn4590868'
ALL_USED_IDs = c(ALL_USED_IDs, UPDRS_ID)
UPDRS_TBL = synTableQuery(paste('SELECT * FROM',UPDRS_ID)) 
UPDRS = UPDRS_TBL@values
```
Voice data with `r dim(VOICE)[1]` recordings from `r length(unique(VOICE$healthCode))` participants are downloaded from `r VOICE_ID`

### Data quality check
`r length(VOICE$recordId[is.na(VOICE$audio_audio.m4a) & is.na(VOICE$audio_countdown.m4a)])` records are missing audio and countdown. 

Their record ids are `r paste(VOICE$recordId[is.na(VOICE$audio_audio.m4a) & is.na(VOICE$audio_countdown.m4a)], collapse = ',')`

`r length(VOICE$recordId[is.na(VOICE$audio_audio.m4a) & !is.na(VOICE$audio_countdown.m4a)])` records are missing audio alone. 

Their ids are `r paste(VOICE$recordId[is.na(VOICE$audio_audio.m4a) & !is.na(VOICE$audio_countdown.m4a)], collapse=',')`

`r length(VOICE$recordId[!is.na(VOICE$audio_audio.m4a) & is.na(VOICE$audio_countdown.m4a)])` records are missing countdown alone. 

Their ids are `r paste(VOICE$recordId[!is.na(VOICE$audio_audio.m4a) & is.na(VOICE$audio_countdown.m4a)], collapse=',')`

### Download audio_audio m4a files
```{r audio.download, include=FALSE, eval=FALSE}
Files = list()
for (i in 1:ceiling(dim(VOICE)/1000)[1]){
  write.table(paste0('Downloading chunk ',i,'...'), file='count.txt',append=T, row.names=F, col.names=F)
  ind.start = (i-1)*1000+1
  ind.end = min(i*1000, dim(VOICE)[1])
  Files[[i]] = getVoiceData(VOICE$rowID[ind.start:ind.end], VOICE_TBL, "audio_audio.m4a")  
}
Files.Name.Mapping = rbindlist(lapply(Files, function(x){tmp = data.frame(rowID = names(x$fnms), fileLocation = x$fnms)}))
write.table(Files.Name.Mapping, file='audioFileMapping.tsv', sep= '\t',row.names=F, quote=F)
```

### Convert audio_audio m4a to wav files
```{r audio.m4a2wav, include=FALSE, eval=FALSE}
Files.Present = list.files(path=synapseCacheDir(), pattern='audio_audio.*.tmp', recursive=T)

convertM4A2WAV <- function(file.name){
  m4a.file = paste(synapseCacheDir(),file.name,sep='/')
  wav.file = paste(tools::file_path_sans_ext(m4a.file),'wav',sep='.')
  
  system(paste('avconv','-i',m4a.file,'-b 64k',wav.file))
}
Files.conv = lapply(Files.Present, convertM4A2WAV)
```

### Unconverted audio_audio m4a files
```{r unconv.file}
Files.Name.Mapping = fread('./audioFileMapping.tsv', data.table=F)
  
Files.m4a.sansExt = sapply(Files.Name.Mapping$fileLocation,tools::file_path_sans_ext)
Files.wav.sansExt = sapply(list.files(path=synapseCacheDir(), pattern='audio_audio.*.wav', recursive=T),
                           function(x){
                             x = tools::file_path_sans_ext(x);
                             x = paste(synapseCacheDir(),x,sep='/');
                             return(x)
                           })
tmp = sapply(setdiff(Files.m4a.sansExt,Files.wav.sansExt), function(x) paste0(x,'.tmp'))
tmp = VOICE[Files.Name.Mapping$rowID[(Files.Name.Mapping$fileLocation %in% tmp)], 'recordId']
```
Following `r length(tmp)` m4a audio_audio files cannot be converted into wav files: 

RecordIDs - `r paste(tmp, collapse = ',')`

### Download audio_countdown m4a files
```{r countdown.download, include=FALSE, eval=FALSE}
Files = list()
for (i in 1:ceiling(dim(VOICE)/1000)[1]){
  write.table(paste0('Downloading chunk ',i,'...'), file='count.txt',append=T, row.names=F, col.names=F)
  ind.start = (i-1)*1000+1
  ind.end = min(i*1000, dim(VOICE)[1])
  Files[[i]] = getVoiceData(VOICE$rowID[ind.start:ind.end], VOICE_TBL, "audio_countdown.m4a")  
}
Files.Name.Mapping = rbindlist(lapply(Files, function(x){tmp = data.frame(rowID = names(x$fnms), fileLocation = x$fnms)}))
write.table(Files.Name.Mapping, file='countdownFileMapping.tsv', sep= '\t',row.names=F, quote=F)
```

### Convert audio_countdown m4a to wav files
```{r countdown.m4a2wav, include=FALSE, eval=FALSE}
Files.Present = list.files(path=synapseCacheDir(), pattern='audio_countdown.*.tmp', recursive=T)

Files.conv = lapply(Files.Present, convertM4A2WAV)
```

### Unconverted audio_countdown m4a files
```{r unconv.countdown.file}
Files.Name.Mapping = fread('./countdownFileMapping.tsv', data.table=F)
  
Files.m4a.sansExt = sapply(Files.Name.Mapping$fileLocation,tools::file_path_sans_ext)
Files.wav.sansExt = sapply(list.files(path=synapseCacheDir(), pattern='audio_countdown.*.wav', recursive=T),
                           function(x){
                             x = tools::file_path_sans_ext(x);
                             x = paste(synapseCacheDir(),x,sep='/');
                             return(x)
                           })
tmp = sapply(setdiff(Files.m4a.sansExt,Files.wav.sansExt), function(x) paste0(x,'.tmp'))
tmp = VOICE[Files.Name.Mapping$rowID[(Files.Name.Mapping$fileLocation %in% tmp)], 'recordId']
```
Following `r length(tmp)` m4a audio_countdown files cannot be converted into wav files: 

RecordIDs - `r paste(tmp, collapse = ',')`

### Download json files
```{r json.download, include=FALSE, eval=FALSE}
Files = list()
for (i in 1:ceiling(dim(VOICE)/1000)[1]){
  write.table(paste0('Downloading chunk ',i,'...'), file='count.txt',append=T, row.names=F, col.names=F)
  ind.start = (i-1)*1000+1
  ind.end = min(i*1000, dim(VOICE)[1])
  Files[[i]] = getVoiceData(VOICE$rowID[ind.start:ind.end], VOICE_TBL, "momentInDayFormat.json")  
}
Files.Name.Mapping = rbindlist(lapply(Files, function(x){tmp = data.frame(rowID = names(x$fnms), fileLocation = x$fnms)}))
write.table(Files.Name.Mapping, file='jsonFileMapping.tsv', sep= '\t',row.names=F, quote=F)
```
```{r extract.json, eval=FALSE}
Files.Name.Mapping = fread('./jsonFileMapping.tsv', data.table=F)
setnames(Files.Name.Mapping, 'fileLocation', 'jsonFileLocation')

choiceAnswers = sapply(Files.Name.Mapping$jsonFileLocation, function(x){return(fromJSON(file = x)$choiceAnswers)})
l = sapply(choiceAnswers, length)
choiceAnswers[l==0] = lapply(choiceAnswers[l==0], function(x) x= NA)
choiceAnswers = ldply(choiceAnswers)
colnames(choiceAnswers) = c('jsonFileLocation','choiceAnswer')

VOICE = unique(merge(VOICE, merge(Files.Name.Mapping, choiceAnswers, by = 'jsonFileLocation', all = T), by = 'rowID', all = T))
```
### Extract voice features from audio and countdown files
Source extraction is performed in matlab with signal processing toolbox and the results are imported here
```{r feature.extraction.audio}
Audio = fread('/mnt/Github/mPower_Data_Analysis/code//Matlab//Audio_Features.csv', data.table = F)
setnames(Audio, 'Row', 'audioFileLocation.wav')
colnames(Audio)[2:14] = paste('audio', colnames(Audio)[2:14], sep = '.')

AudioFileNames = fread('./audioFileMapping.tsv', data.table=F)
setnames(AudioFileNames, 'fileLocation', 'audioFileLocation.m4a')
AudioFileNames = dplyr::mutate(AudioFileNames, audioFileLocation.wav = gsub('.tmp','.wav',audioFileLocation.m4a))

# Merge file name mapping and features
VOICE = unique(merge(VOICE, merge(AudioFileNames, Audio, by='audioFileLocation.wav', all.x = T), by = 'rowID', all = T))
```
```{r feature.extraction.countdown}
Countdown = fread('/mnt/Github/mPower_Data_Analysis/code/Matlab//Countdown_Features.csv', data.table=F)
setnames(Countdown, 'Row', 'countdownFileLocation.wav')
colnames(Countdown)[2:14] = paste('countdown', colnames(Countdown)[2:14], sep = '.')

CountdownFileNames = fread('./countdownFileMapping.tsv', data.table=F)
setnames(CountdownFileNames, 'fileLocation', 'countdownFileLocation.m4a')
CountdownFileNames = dplyr::mutate(CountdownFileNames, countdownFileLocation.wav = gsub('.tmp','.wav',countdownFileLocation.m4a))

# Merge file name mapping and features
VOICE = unique(merge(VOICE, merge( CountdownFileNames, Countdown, by='countdownFileLocation.wav', all.x = T), by = 'rowID', all = T))
```
### Extract survey data
```{r extract.survey}
FactCovars = c("healthCode","appVersion","phoneInfo","are-caretaker","deep-brain-stimulation","diagnosis-year","education","employment","gender","healthcare-provider","maritalStatus","medical-usage","medical-usage-yesterday","onset-year","professional-diagnosis","race","smartphone","smoked","surgery")
ContCovars = c("age","years-smoking")
DateVars = c("uploadDate","createdOn","last-smoked")
SURVEY = SURVEY[,c(FactCovars, ContCovars, DateVars)]

# Replace all alpha numeric characters in the metadata
myFix1 <- function(x) str_replace_all(x, '[^[:alnum:]]', '')

# Substitute empty cells with NA 
myFix2 <- function(x){ 
  z <- gsub("\\s+", "", x)  #make sure it's "" and not " " etc
  x[z==""] <- NA 
  return(x)
}

SURVEY[,FactCovars] <- SURVEY[,FactCovars] %>%
  dplyr::mutate_each(funs(myFix1), -healthCode, -appVersion, -phoneInfo) %>%
  dplyr::mutate_each(funs(myFix2), -healthCode, -appVersion, -phoneInfo) 

# Convert factor covariates to factors
SURVEY[,FactCovars] = lapply(SURVEY[,FactCovars], factor)
SURVEY[,ContCovars] = lapply(SURVEY[,ContCovars], as.numeric)
SURVEY[,DateVars] = lapply(SURVEY[,DateVars], as.POSIXct)

SURVEY <- SURVEY[!duplicated(SURVEY$healthCode),]

# Find voice data without any recordIds
No.health.record = setdiff(unique(VOICE$healthCode),unique(SURVEY$healthCode))
```
`r length(No.health.record)` voice recordings doesn't have matching metadata

### Extract UPDRS data
```{r extract.UPDRS}
FactCovars = c("healthCode", "appVersion", "phoneInfo", "EQ-5D1", "GELTQ-1a", "GELTQ-1b", "GELTQ-1c", "GELTQ-2", "MDS-UPRDRS1.1", "MDS-UPDRS1.3",
               "MDS-UPDRS1.4", "MDS-UPDRS1.5","MDS-UPDRS1.7","MDS-UPDRS1.8","MDS-UPDRS2.1", "MDS-UPDRS2.4", "MDS-UPDRS2.5", "MDS-UPDRS2.6", 
               "MDS-UPDRS2.7", "MDS-UPDRS2.8", "MDS-UPDRS2.9", "MDS-UPDRS2.10", "MDS-UPDRS2.12", "MDS-UPDRS2.13")
DateVars = c("uploadDate","createdOn")
UPDRS = UPDRS[,c(FactCovars, DateVars)]

# Replace all special characters in UPDRS
UPDRS[,FactCovars] <- UPDRS[,FactCovars] %>%
  dplyr::mutate_each(funs(myFix1), -healthCode, -appVersion, -phoneInfo) %>%
  dplyr::mutate_each(funs(myFix2), -healthCode, -appVersion, -phoneInfo) 

# Convert factor covariates to factors
UPDRS[,FactCovars] = lapply(UPDRS[,FactCovars], factor)
UPDRS[,DateVars] = lapply(UPDRS[,DateVars], as.POSIXct)

# Find voice data without any recordIds
No.UPDRS = setdiff(unique(VOICE$healthCode),unique(UPDRS$healthCode))
```
Following `r length(No.UPDRS)` voice recordings doesn't have matching UPDRS

### Store data in synapse
```{r synapse.store, include=FALSE}
# Synapse store parameters
projectId <- "syn4557550"
parentId <- projectId

ActivityName <- 'Feature extraction and quality control analysis of voice data'
  
thisFileName <- 'VoiceDataDownload.Rmd'
  
# Github link
thisRepo <- getRepo(repository = "th1vairam/mPower_Data_Analysis", 
                    ref="branch", 
                    refName='master')

thisFile <- getPermlink(repository = thisRepo,
                        repositoryPath=paste0('code/Rmd/', thisFileName))

# Code
CODE <- File('./VoiceDataDownload.Rmd',name = 'Phonation Feature Extraction and Quality Control',parentId = parentId)
CODE <- synStore(CODE, used = ALL_USED_IDs, activityName = ActivityName, executed = thisFile)

# Data file
write.table(VOICE, file = './Phonation_Features.tsv', sep = '\t', row.names=F, quote=T)
VOICE_OBJ <- File('./Phonation_Features.tsv',name = 'Phonation',parentId = parentId)
VOICE_OBJ <- synStore(VOICE_OBJ, used = ALL_USED_IDs, activityName = ActivityName, executed = thisFile)

# Data file
write.table(SURVEY, file = './SURVEY.tsv', sep = '\t', row.names=F, quote=T)
SUR_OBJ <- File('./SURVEY.tsv',name = 'Survey',parentId = parentId)
SUR_OBJ <- synStore(SUR_OBJ, used = ALL_USED_IDs, activityName = ActivityName, executed = thisFile)

# Data file
write.table(UPDRS, file = './UPDRS.tsv', sep = '\t', row.names=F, quote=T)
UPDRS_OBJ <- File('./UPDRS.tsv',name = 'UPDRS',parentId = parentId)
UPDRS_OBJ <- synStore(UPDRS_OBJ, used = ALL_USED_IDs, activityName = ActivityName, executed = thisFile)
```
|  *Results*           |  *SynapseID*                  |
|  -------             |   ---------                   |
|  VOICE DATA          |  `r paste(VOICE_OBJ$properties$id,VOICE_OBJ$properties$versionNumber, sep ='.')`                  |
|  SURVEY DATA         |  `r paste(SUR_OBJ$properties$id,SUR_OBJ$properties$versionNumber,sep ='.')`                       |
|  UPDRS DATA          |  `r paste(UPDRS_OBJ$properties$id,UPDRS_OBJ$properties$versionNumber, sep ='.')`                  |